---
title: Descriptive Analysis of network graph characteristics
---

Questions about structures or characteristics of a network graph:
* Paths of graphs, flows of commodities
* important nodes as centrality
* groups of similar vertices as graph partitioning problem 

```{r libraries}
library(sand)
library(igraph)
library(ggplot2)
```

Vertex and Edge characteristics

1. Vertex Degree
   In a network of G = (V, E), the degree *d~v~* of a vertex *v* is the number of edges incident upon *v*   
   Given G, *f~d~* is the fraction of vertices *v* with degree *d~v~* = d
   The collection {*f~d~*}~d>=0~ is called the *degree distribution of G*; it's a rescaling of the set of degree frequencies from the original degree sequence
   For weighted networks, a generalization of degree is the notion of vertex *strength*, obtained by summing up all the weights of edges incident to the given vertex; its distribution is the *weighted degree distribution*, 

```{r karate}
data(karate)
f_d  <- degree(karate)
plot(f_d)
hist(f_d, col="cornflowerblue", xlim=c(0,50), xlab="Vertex Degree", ylab = "Frequency", main = "")
w <- strength(karate)
hist(w, col="pink", xlim=c(0,50), xlab="Vertex Strength", ylab = "Frequency", main = "")

```


Shapes of degree distribution can vary greatly

```{r yeast}
data(yeast)
d.yeast <- degree(yeast)
w.yeast <- strength(yeast)
hist(d.yeast, col="cornflowerblue", xlim=c(0,50), xlab="Vertex Degree", ylab = "Frequency", main = "")
hist(w.yeast, col="cornflowerblue", xlim=c(0,50), xlab="Vertex Degree", ylab = "Frequency", main = "")
```

Creating a log-scale for this distribution:

```{r log_scale}
dd.yeast <- degree_distribution(yeast)
d <- 1:max(d.yeast) -1
ind <- (dd.yeast != 0 )
plot(d[ind], dd.yeast[ind], log = "xy", col="blue", xlab=c('Log-Degree'), ylab=c('Log-Intensity'), main="Log-Log Degree Distribution")
```

The average degree of the *neighbors* of a vertex
* e.g. vertex degree as a function of average vertex degree of neighbors
* are higher degree vertices tend to connect to similar degree neighbors?

```{r avg. neighbor degree}
a.nn.deg.yeast <- knn(yeast, V(yeast))$knn
length(a.nn.deg.yeast)

plot(d.yeast, a.nn.deg.yeast, log = "xy", col="goldenrod", xlab=c('Log Vertex Degree'), ylab=c('Log Average Neighbor Degree'))

ggplot() +
   geom_point(aes(x = log(d.yeast), y =log(a.nn.deg.yeast)), color = "goldenrod")
```

## Vertex Centrality
* vertex importance
* one such measure is 'vertex degree'

1. closeness centrality
   * a vertex is central if it's close to many others
   * centrality varies inversely with a measure of total distance of a vertex from all others,

     $$c_{Cl} (\textit{v}) = \frac{1}{\sum_{u \in V} dist(\textit{v}, u)}$$
   * dist(v, u) is the geodesic distance between the vertices $\textit{u, v} \in V$
   * often closeness is normalized to interval [0,1] by multiplication by a factor $ $\textit{N_v} - 1$ $


2. betweenness centrality
   * summarize the extent a vertex is located 'between' other pairs of vertices
   * in this sense importance means where a vertex is located with respect to the paths of the network
   * vertices that sit on many paths are considered important
   * Freeman's formula

   \[ c_{B}(V) = \sum_{s \neq t \neq v \in V} \frac {\sigma (s, t \mid v)} {\sigma(s, t)} \]

   where the nominator is the total number of shortest paths between s and t, which pass through v; the denominator is the total number of shortest paths between s and t.
   * mostly rescaled by division of a factor of (N_v -1) (N_v -2) /2


3. eigenvector centrality
   * the notion of 'status', 'rank' or 'prestige'
   * the more central the neighbors are, the more central the vertex is
   * can be defined as eigenvector solutions of linear systems of equations
   * there are many eigenvector centrality measures, such as Bonacich's:

   \[ c_{E_i}(v) = \alpha \sum _{[u, v] \in E} C_{E_i}(u) \]






